---
title: "Test report"
author: "M. Salmon"
date: "January 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

# Here is the first section

## Here is the first sub-section

I want to put *these words* in bold and _these ones_ in italic.

I want this command to look like one `summary(cars)`.

I also want to write a list

- Element 1

- Element 2

# Now I want to show a table but in a pretty way
```{r ugly table}
head(cars)

```


```{r pretty table}
knitr::kable(head(cars) )

```

What if I want to write in the text the size of my data? Better not to copy-paste. There are `r nrow(cars)` entries in my dataset.

## Regression

```{r regression}
model <- lm(speed ~ dist, data = cars)
model$coefficients
library("broom")
tidy(model)
knitr::kable(tidy(model))

summary(model)
knitr::kable(glance(model))

``` 

## Figure options

```{r fig1, fig.width = 5}

plot(cars)
```


```{r fig2, fig.width = 10}

plot(cars)
```